---
title: "Manuscript Figures & Tables"
subtitle: "Background"
author: "J.P. Meagher"
format:
  pdf:
    keep-tex: true
    fig-width: 6
    fig-crop: false
execute:
  echo: false
  warning: false
  message: false
knitr:
  opts_chunk:
    crop: null
    dev: "pdf"
---

```{r setup}
library(onlineMessageboardActivity)
library(dplyr)
library(ggplot2)
library(latex2exp)
library(lubridate)
library(magrittr)
library(zoo)
library(ggpubr)

source(here::here("code_modules/plotting.R"))
```

# Exploratory analysis

This section of the manuscript includes three figures exploring various aspects of the `r/ireland` dataset.

## Discussion structure and daily activity

This figure highlights the branching structure associated with a typical sequence of discussions and the circadian rhythm associated with overall activity on the subreddit.

```{r data_snapshot_plot}
# Select an interval to examine (cherry-picked for illustration)
lower <- ymd_hms(20190415160000)
upper <- lower + minutes(30)

# Identify the discussions in the interval
discussions <- messageboard_df |>
  filter(parent_id == 0) |>
  filter(time > lower & time < upper) |>
  use_series(discussion)

# Pull out and prepare the subset of discussions
subset <- messageboard_df |>
  filter(discussion %in% discussions) |>
  mutate(
    parent_id = refactor_branching_structure(
      id = id,
      parent_id = parent_id,
      is_immigrant = (parent_id == 0),
      S = 20
    )
  ) |>
  mutate(id = row_number()) |>
  mutate(
    t = as.numeric(
      difftime(
        time,
        dmy_hms(010419000000, tz = "Europe/London"),
        units = "hours"
      )
    )
  ) |>
  mutate(t = t - min(t)) |>
  group_by(discussion) |>
  mutate(tau = t - min(t)) |>
  ungroup() |>
  mutate(type = parent_id != 0) |>
  mutate(type = factor(type, labels = c("post", "comment")))

# Specify the generation of each node
generation <- rep(NA, nrow(subset))
generation[subset$parent_id == 0] <- 0
for (i in 1:nrow(subset)) {
  if (is.na(generation[i])) {
    generation[i] <- generation[subset$parent_id[i]] + 1
  }
}

subset <- subset |>
  mutate(generation = generation) |>
  mutate(generation = factor(generation, ordered = TRUE)) |>
  mutate(
    n_offspring = tabulate(parent_id + 1, nbins = nrow(subset) + 1)[-1]
  )

# Set the origin for each edge in the branching structure
# Perturb child nodes along the x-axis for clearer plotting
new_v <- subset$parent_id
perturb <- source_x <- source_y <- off_ind <- rep(0, nrow(subset))
per_range <- 0.05

for (i in 1:nrow(subset)) {
  if (subset$parent_id[i] != 0) {
    off_ind[new_v[i]] <- off_ind[new_v[i]] + 1
    if (subset$generation[i] == 1) {
      perturb[i] <- seq(
        -per_range,
        per_range,
        length.out = subset$n_offspring[new_v[i]] + 2
      )[off_ind[new_v[i]] + 1]
    } else {
      perturb[i] <- perturb[new_v[i]] +
        seq(
          -per_range,
          per_range,
          length.out = subset$n_offspring[new_v[i]] + 2
        )[off_ind[new_v[i]] + 1]
    }
    source_x[i] <- (
      subset$time -
        (seconds((subset$tau + perturb) * 60 * 60))
    )[new_v[i]]
    source_y[i] <- subset$tau[new_v[i]]
  } else {
    source_x[i] <- subset$time[i]
    source_y[i] <- NA
  }
}

# Convert x axis to clock time
subset$perturb <- perturb
subset$source_x <- as.POSIXct(
  source_x,
  origin = "1970-01-01",
  tz = "Europe/London"
)
subset$source_y <- source_y

# Make plot
gg_snapshot <- ggplot(subset) +
  geom_segment(
    aes(
      x = time - seconds((tau + perturb) * 60 * 60),
      y = tau + 1,
      xend = source_x,
      yend = source_y + 1
    )
  ) +
  geom_point(
    aes(
      x = time - seconds((tau + perturb) * 60 * 60),
      y = tau + 1,
      color = generation
    )
  ) +
  my_theme +
  theme(legend.position = "none") +
  scale_y_log10(breaks = 2^(0:5), labels = 2^(0:5) - 1) +
  scale_color_viridis_d() +
  labs(
    title = "The structure of discussion trees on r/ireland",
    x = "Monday, April 15, 2019",
    y = TeX("Hours since post"),
    color = NULL
  )
```

```{r activity_levels_plot}
n_buckets <- 5
window <- 3

gg_activity <- messageboard_df |>
  mutate(moy = month(time)) |>
  mutate(dow = wday(time, week_start = 1)) |>
  mutate(dow = factor(dow)) |>
  mutate(dom = day(time)) |>
  mutate(hod = hour(time)) |>
  mutate(moh = minute(time)) |>
  mutate(
    bucket = sapply(
      moh, function(x) sum(x >= seq(0, 60, length.out = n_buckets + 1))
    )
  ) |>
  group_by(moy, dow, dom, hod, bucket) |>
  summarise(count = max(0, length(id)), .groups = "drop") |>
  mutate(t = hod + (bucket - 0.5) / n_buckets) |>
  mutate(date = paste0(dom, "-", moy)) |>
  mutate(
    ma_count = n_buckets * rollmean(count, window * n_buckets, na.pad = TRUE)
  ) |>
  ggplot() +
  geom_line(
    aes(x = t, y = ma_count, group = date),
    color = "#440154",
    alpha = 0.5
  ) +
  my_theme +
  theme(legend.position = "none") +
  scale_color_viridis_d() +
  scale_x_continuous(breaks = c(0, 6, 12, 18, 24)) +
  labs(
    title = TeX("Daily activity levels on r/ireland"),
    subtitle = paste0(window, "-hour moving average"),
    x = "Hour of the day (IST)",
    y = "Nodes per hour"
  )
```

```{r 02_background_fig-data_exploration, fig.height=4, fig.width=6}
ggarrange(
  gg_snapshot,
  gg_activity,
  labels = "auto",
  nrow = 2
)
```

## Reply counts and overdispersion

This figure highlights that the mean number of replies to each point within the subreddit seems to follow a circadian rhythm and that the distribution of the number of replies to each point is overdispersed relative to the Poisson distribution.

```{r mean_replies}
n_buckets <- 5
window <- 3

gg_mean_replies <- messageboard_df |>
  mutate(
    z = tabulate(parent_id + 1, nbins = nrow(messageboard_df) + 1)[-1]
  ) |>
  mutate(moy = month(time)) |>
  mutate(dow = wday(time, week_start = 1)) |>
  mutate(dow = factor(dow)) |>
  mutate(dom = day(time)) |>
  mutate(hod = hour(time)) |>
  mutate(moh = minute(time)) |>
  mutate(
    bucket = sapply(
      moh,
      function(x) sum(x >= seq(0, 60, length.out = n_buckets + 1))
    )
  ) |>
  filter(parent_id == 0) |>
  group_by(moy, dow, dom, hod, bucket) |>
  summarise(replies = mean(z), .groups = "drop") |>
  mutate(t = hod + (bucket - 0.5) / n_buckets) |>
  mutate(date = paste0(dom, "-", moy)) |>
  mutate(
    ma_replies = rollmean(replies, window * n_buckets, na.pad = TRUE)
  ) |>
  group_by(t) |>
  summarise(mean_replies = mean(ma_replies, na.rm = TRUE), .groups = "drop") |>
  cbind(type = "post") |>
  rbind(
    messageboard_df |>
      mutate(
        z = tabulate(parent_id + 1, nbins = nrow(messageboard_df) + 1)[-1]
      ) |>
      mutate(moy = month(time)) |>
      mutate(dow = wday(time, week_start = 1)) |>
      mutate(dow = factor(dow)) |>
      mutate(dom = day(time)) |>
      mutate(hod = hour(time)) |>
      mutate(moh = minute(time)) |>
      mutate(
        bucket = sapply(
          moh,
          function(x) sum(x >= seq(0, 60, length.out = n_buckets + 1))
        )
      ) |>
      filter(parent_id != 0) |>
      group_by(moy, dow, dom, hod, bucket) |>
      summarise(replies = mean(z), .groups = "drop") |>
      mutate(t = hod + (bucket - 0.5) / n_buckets) |>
      mutate(date = paste0(dom, "-", moy)) |>
      mutate(
        ma_replies = rollmean(replies, window * n_buckets, na.pad = TRUE)
      ) |>
      group_by(t) |>
      summarise(mean_replies = mean(ma_replies, na.rm = TRUE), .groups = "drop") |>
      cbind(type = "comment")
  ) |>
  ggplot() +
  geom_line(
    aes(x = t, y = mean_replies, color = type)
  ) +
  my_theme +
  theme(legend.position = "bottom") +
  scale_x_continuous(breaks = c(0, 6, 12, 18, 24)) +
  scale_color_viridis_d() +
  labs(
    title = TeX("Replies on r/ireland"),
    subtitle = paste0(window, "-hour moving average"),
    x = "Hour of the day (IST)",
    y = "Number of replies",
    color = NULL
  )
```

```{r reply_distribution}
R <- messageboard_df |>
  mutate(z = tabulate(parent_id + 1, nbins = nrow(messageboard_df) + 1)[-1]) |>
  use_series(z) |>
  mean()

posts_z <- messageboard_df |>
  mutate(z = tabulate(parent_id + 1, nbins = nrow(messageboard_df) + 1)[-1]) |>
  filter(parent_id == 0) |>
  use_series(z) |>
  tabulate()

comments_z <- messageboard_df |>
  mutate(z = tabulate(parent_id + 1, nbins = nrow(messageboard_df) + 1)[-1]) |>
  filter(parent_id != 0) |>
  use_series(z) |>
  tabulate()

gg_reply_distribution <- left_join(
  data.frame(Posts = posts_z / sum(posts_z)) |>
    mutate(n = -1 + row_number()) |>
    filter(Posts != 0),
  data.frame(Comments = comments_z / sum(comments_z)) |>
    mutate(n = -1 + row_number()) |>
    filter(Comments != 0),
  by = "n"
) |>
  reshape2::melt(id.vars = "n") |>
  ggplot() +
  geom_line(
    data = data.frame(x = 0:9, y = dpois(0:9, lambda = R)),
    aes(x = x, y = y),
    lty = 2
  ) +
  geom_point(
    aes(x = n, y = value, color = variable)
  ) +
  scale_color_viridis_d() +
  scale_fill_viridis_d() +
  scale_y_log10() +
  my_theme +
  theme(legend.position = "top") +
  labs(
    title = TeX("Reply count distribution"),
    x = "Number of replies",
    y = "Density (log scale)",
    fill = NULL,
    color = NULL
  )
```

```{r 02_background_fig-reply_distribution_exploration, fig.height=3, fig.width=6}
ggarrange(
  gg_mean_replies, gg_reply_distribution,
  nrow = 1, labels = "auto",
  common.legend = TRUE, legend = "bottom"
)
```

## Time-of-day dependence in generation intervals

The distribution of generation intervals depends on the time of day that a point arrives.

```{r first_reply_interval}
gg_reply_interval_0_2 <- messageboard_df |>
  mutate(
    t = as.numeric(
      difftime(
        time,
        dmy_hms(010419000000, tz = "Europe/London"),
        units = "hours"
      )
    )
  ) |>
  mutate(
    parent_t = sapply(
      parent_id,
      function(i) {
        if (i == 0) NA else t[i]
      }
    )
  ) |>
  mutate(
    type = sapply(
      parent_id,
      function(i) {
        if (i == 0) {
          NA
        } else if (parent_id[i] == 0) {
          "post"
        } else {
          "comment"
        }
      }
    )
  ) |>
  mutate(tau = t - parent_t) |>
  mutate(parent_tod = mod(parent_t, 24)) |>
  mutate(parent_tod = floor(parent_tod)) |>
  filter(parent_tod > 0 & parent_tod < 2) |>
  filter(tau < 24) |>
  ggplot() +
  geom_histogram(
    aes(x = tau, y = after_stat(density), fill = type),
    alpha = 0.5,
    position = "identity",
    binwidth = 0.5,
    boundary = 0
  ) +
  scale_fill_viridis_d() +
  scale_color_viridis_d() +
  my_theme +
  theme(legend.position = "top") +
  scale_x_continuous(breaks = c(0, 6, 12, 18, 24)) +
  labs(
    title = "Reply interval distribution",
    subtitle = "Parent nodes during 00:00-02:00",
    x = TeX("Hours since parent node"),
    y = "Density",
    fill = NULL,
    color = NULL
  )
```

```{r second_reply_interval}
gg_reply_interval_12_14 <- messageboard_df |>
  mutate(
    t = as.numeric(
      difftime(
        time,
        dmy_hms(010419000000, tz = "Europe/London"),
        units = "hours"
      )
    )
  ) |>
  mutate(
    parent_t = sapply(
      parent_id,
      function(i) {
        if (i == 0) NA else t[i]
      }
    )
  ) |>
  mutate(
    type = sapply(
      parent_id,
      function(i) {
        if (i == 0) {
          NA
        } else if (parent_id[i] == 0) {
          "post"
        } else {
          "comment"
        }
      }
    )
  ) |>
  mutate(tau = t - parent_t) |>
  mutate(parent_tod = mod(parent_t, 24)) |>
  mutate(parent_tod = floor(parent_tod)) |>
  filter(parent_tod > 12 & parent_tod < 14) |>
  filter(tau < 24) |>
  ggplot() +
  geom_histogram(
    aes(x = tau, y = after_stat(density), fill = type),
    alpha = 0.5,
    position = "identity",
    binwidth = 0.5,
    boundary = 0
  ) +
  scale_fill_viridis_d() +
  scale_color_viridis_d() +
  my_theme +
  theme(legend.position = "top") +
  scale_x_continuous(breaks = c(0, 6, 12, 18, 24)) +
  labs(
    title = "Reply interval distribution",
    subtitle = "Parent nodes during 12:00-14:00",
    x = TeX("Hours since parent node"),
    y = "Density",
    fill = NULL,
    color = NULL
  )
```

```{r 02_background_fig-reply_interval_exploration, fig.height=3, fig.width=6}
ggarrange(
  gg_reply_interval_0_2, gg_reply_interval_12_14,
  nrow = 1, labels = "auto",
  common.legend = TRUE, legend = "bottom"
)
```
