---
title: "Manuscript Figures & Tables"
subtitle: "Methods"
author: "J.P. Meagher"
format:
  pdf:
    keep-tex: true
    fig-width: 6
    fig-crop: false
execute:
  echo: false
  warning: false
  message: false
knitr:
  opts_chunk:
    crop: null
    dev: "pdf"
---

```{r setup}
library(onlineMessageboardActivity)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(lubridate)
library(latex2exp)

source(here::here("code_modules/plotting.R"))
```

This section of the manuscript presents figures illustrating the generative model.

## Discussion tree structure

These figures illustrate the branching structure for a single discussion tree on r/ireland.

```{r identify_discussion_tree}
# Plot discussion 14491 as an illustrative example
# Chosen as it sits in snapshot interval in the Background section
# and illustrates the branching structure well.
discussion_id <- 14491

discussion_tree <- messageboard_df |>
  filter(discussion == discussion_id) |>
  mutate(
    parent_id = refactor_branching_structure(
      id = id,
      parent_id = parent_id,
      is_immigrant = (parent_id == 0),
      S = 1
    )
  ) |>
  mutate(id = seq_along(id)) |>
  mutate(generation = ifelse(parent_id == 0, 0, NA))

# Compute generation for each node (depth in tree)
for (i in 1:nrow(discussion_tree)) {
  if (is.na(discussion_tree$generation[i])) {
    discussion_tree$generation[i] <-
      discussion_tree$generation[discussion_tree$parent_id[i]] + 1
  }
}
```

```{r construct_objects_for_discussion_tree_plot}
# Manually position nodes on y-axis to avoid overlapping branches
plot_df <- discussion_tree |>
  mutate(
    manual_y_position = c(0, 0.5, 1, 1.5, 2, 0.5, 1, 2.25, 1.75, 2.5, 1.5)
  )

# Create line segment data for connecting parent-child nodes
segments_df <- plot_df |>
  filter(parent_id != 0) |>
  rowwise() |>
  mutate(
    parent_time = plot_df$time[plot_df$id == parent_id],
    parent_y = plot_df$manual_y_position[plot_df$id == parent_id]
  ) |>
  ungroup()
```

```{r}
#| label: fig-discussion-tree
#| fig-cap: "The branching structure of a discussion tree on the r/ireland subreddit. The initial post and subsequent comments are modelled as sequence of points. Line segments represent the reply-to relations between points. Each point is indexed such that we have $\\boldsymbol y_{t_n} = \\left(\\boldsymbol t, \\boldsymbol \\beta\\right)$ where $n = 11$, $\\boldsymbol t = \\left(t_1, t_2, \\dots t_{11}\\right)$ are timestamps, and $\\boldsymbol \\beta = \\left( \\beta_1, \\beta_2, \\dots, \\beta_{11}\\right)$ links each point to its parent witing in the branching structure. For example, $\\beta_5 = 1$, $\\beta_8 = 5$ and $\\beta_{11} = 7$."
#| fig-height: 2.5

individual_tree_plot <-
  ggplot() +
  geom_segment(
    data = segments_df,
    aes(
      x = time, xend = time,
      y = parent_y, yend = manual_y_position
    ),
    color = "black"
  ) +
  geom_segment(
    data = segments_df,
    aes(
      x = parent_time, xend = time,
      y = parent_y, yend = parent_y
    ),
    color = "black"
  ) +
  geom_point(
    data = plot_df,
    aes(x = time, y = manual_y_position, color = as.factor(generation)),
    size = 3
  ) +
  geom_text(
    data = plot_df,
    aes(x = time, y = manual_y_position, label = id),
    hjust = 1.5,
    vjust = -1,
    size = 3,
    fontface = "bold"
  ) +
  coord_cartesian(ylim = c(-0.5, 3.5)) +
  my_theme +
  theme(
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank(),
    axis.line.y = element_blank(),
    legend.position = "bottom"
  ) +
  scale_color_viridis_d() +
  labs(
    title = "The structure of a single discussion tree on r/ireland",
    x = "Monday, April 15, 2019",
    y = NULL,
    colour = "Generation"
  )

individual_tree_plot
```

```{r construct_objects_for_alt_tree_plot}
alt_plot_df <- discussion_tree |>
  mutate(
    manual_y_position = c(0, 1, -1, 2, -2, -1, 1, -1.5, -2, -2.5, 1),
    node_hjust = c(2, -1, rep(-0.75, 9)),
    node_vjust = c(0.5, 0.5, rep(1.25, 9))
  )

alt_segments_df <- alt_plot_df |>
  filter(parent_id != 0) |>
  rowwise() |>
  mutate(
    parent_time = alt_plot_df$time[alt_plot_df$id == parent_id],
    parent_y = alt_plot_df$manual_y_position[alt_plot_df$id == parent_id]
  ) |>
  ungroup() |>
  mutate(
    midpoint_time = as.POSIXct(
      (as.numeric(parent_time) + as.numeric(time)) / 2,
      origin = "1970-01-01"
    ),
    midpoint_y = (parent_y + manual_y_position) / 2,
    beta_label = paste0("beta[", id, "] == ", parent_id),
    beta_hjust = c(
      0.5 + 0.6, 0.5 + 0.6, 0.5, 0.5 - 0.05, 0.5 - 0.25,
      0.5 - 0.25, 0.5 + 0.5, 0.5 - 1, 0.5, 0.5
    ),
    beta_vjust = c(
      0.5 - 0.5, 0.5 + 0.5, 0.5 - 1.25, 0.5 + 1.25, 0.5 - 1,
      0.5 - 1, 0.5 - 0.5, 0.5 - 0.75, 0.5 + 1, 0.5 - 0.75
    ),
    beta_angle = rep(0, 10)
  )
```

```{r}
#| label: 03_methods_fig-discussion_tree
#| fig-cap: "In our generative model for discussion trees, the initial post and subsequent comments are modelled as sequence of points. Line segments represent the reply-to relations between points. Each point is indexed such that we have the vector of timestamps $\\boldsymbol t = \\left(t_1, \\dots t_{n}\\right)$ and branching structure $\\boldsymbol \\beta = \\left( \\beta_1, \\dots, \\beta_{n}\\right)$ which define the cluster $y_{t_n} = \\left( \\boldsymbol t, \\boldsymbol \\beta \\right)$. In this example, we have $n = 11$, points are indexed in order of their arrival, we have labelled each point with it's index and we have labelled each branch with the relevant entry in the branching structure."
#| fig-height: 4.5

alt_individual_tree_plot <-
  ggplot() +
  geom_segment(
    data = alt_segments_df,
    aes(
      x = parent_time, xend = time,
      y = parent_y, yend = manual_y_position
    ),
    color = "black"
  ) +
  geom_text(
    data = alt_segments_df,
    aes(
      x = midpoint_time,
      y = midpoint_y,
      label = beta_label,
      hjust = beta_hjust,
      vjust = beta_vjust,
      angle = beta_angle
    ),
    size = 5,
    parse = TRUE
  ) +
  geom_point(
    data = alt_plot_df,
    aes(x = time, y = manual_y_position, color = as.factor(generation)),
    size = 5
  ) +
  geom_text(
    data = alt_plot_df,
    aes(
      x = time,
      y = manual_y_position,
      label = paste0("t[", id, "]"),
      hjust = node_hjust,
      vjust = node_vjust
    ),
    size = 5,
    parse = TRUE
  ) +
  coord_cartesian(
    ylim = c(-3, 3),
    xlim = c(ymd_hm("2019-04-15 16:00"), ymd_hm("2019-04-15 22:45"))
  ) +
  my_theme +
  theme(
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank(),
    axis.line.y = element_blank(),
    legend.position = "bottom"
  ) +
  scale_color_viridis_d() +
  labs(
    title = "The structure of a single discussion tree on the r/ireland subreddit",
    x = "Monday, April 15, 2019",
    y = NULL,
    colour = "Generation"
  )

alt_individual_tree_plot
```

\newpage

## Simulated cluster and model components

This figure presents a synthetic cluster, highlighting relevant features of the model. The figure consists of 4 elements.

```{r simulate_cluster}
# Set hyperparameters
day <- 24
f <- 1 / c(day, day / 2)
omega <- 2 * pi * f
alpha <- c(-0.34, -0.47, -0.18, 0.34)

# Simulate a cluster
set.seed(108)
simulated_cluster <- simulate_gpm_cluster_process(
  t_seed = 0, branching_structure_seed = 0,
  observation_horizon = 0, simulation_horizon = 48,
  immigrant_reproduction_number = 4,
  immigrant_gi_exp_decay_rate = 0.1,
  immigrant_dispersion_parameter = 1,
  offspring_reproduction_number = 2,
  offspring_gi_exp_decay_rate = 0.25,
  offspring_dispersion_parameter = 0.25,
  sinusoid_coefficients = alpha,
  sinusoid_frequencies = omega
)
```

```{r aesthetic_adjustment}
# Minor adjustment of one point to improve the aesthetic of the figure
simulated_cluster$t[7] <- simulated_cluster$t[7] + 1.5
```

```{r simulated_cluster_plot}
gg_sim_clust <- simulated_cluster |>
  mutate(type = ifelse(parent_id == 0, "post", "comment")) |>
  mutate(
    y_pos = c(
      -0.025, -0.025, -0.15, -0.1, -0.05,
      0, 0.01, -0.15, -0.075, 0.05,
      -0.09, -0.025, 0.1, 0.025, 0.025,
      -0.025, 0.175, 0.075, 0.18
    )
  ) |>
  mutate(xend = c(NA, t[parent_id])) |>
  mutate(yend = c(NA, y_pos[parent_id])) |>
  ggplot() +
  geom_segment(aes(x = t, y = y_pos, xend = xend, yend = yend)) +
  geom_point(aes(x = t, y = y_pos, size = nu, color = type)) +
  my_theme +
  theme(
    legend.position = "none",
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.line.y = element_blank(),
    axis.ticks.y = element_blank()
  ) +
  scale_x_continuous(breaks = (0:8) * 6, limits = c(0, 48)) +
  scale_color_viridis_d() +
  labs(
    title = "A simulated cluster process",
    x = NULL
  )
```

```{r circadian_rhythm_plot}
gg_activity <- data.frame(t = seq(0, 48, length.out = 1001)) |>
  mutate(alpha = 1 + sinusoidal_function(t, alpha, omega)) |>
  ggplot() +
  geom_line(aes(x = t, y = alpha)) +
  my_theme +
  scale_x_continuous(breaks = (0:8) * 6) +
  labs(
    title = "Activity function",
    x = NULL,
    y = TeX("\\alpha (t)")
  )
```

```{r cumulative_offspring_density_plot}
gg_excitation <- data.frame(t = seq(0, 48, length.out = 1001)) |>
  mutate(
    excitation = sapply(
      t,
      function(x) {
        sum(dexp(
          x - simulated_cluster$t,
          rate = c(0.1, rep(0.25, 18))
        ))
      }
    )
  ) |>
  ggplot() +
  geom_line(aes(x = t, y = excitation)) +
  my_theme +
  scale_x_continuous(breaks = (0:8) * 6) +
  labs(
    title = "Cumulative offspring density",
    x = NULL,
    y = TeX("$\\sum_{t_j < t} \\rho (t - t_j | \\beta_j)$")
  )
```

```{r ground_intensity_process_plot}
gg_intensity <- data.frame(t = seq(0, 48, length.out = 1001)) |>
  mutate(alpha = 1 + sinusoidal_function(t, alpha, omega)) |>
  mutate(
    excitation = sapply(
      t,
      function(x) {
        sum(simulated_cluster$nu * dexp(
          x - simulated_cluster$t,
          rate = c(0.1, rep(0.25, 18))
        ))
      }
    )
  ) |>
  mutate(intensity = alpha * excitation) |>
  ggplot() +
  geom_line(aes(x = t, y = intensity)) +
  my_theme +
  scale_x_continuous(breaks = (0:8) * 6) +
  labs(
    title = "Ground intensity process",
    x = "t",
    y = TeX("$\\lambda^*$(t)")
  )
```

```{r 03_methods_fig-model_form, fig.height=6, fig.width=6}
ggarrange(
  gg_sim_clust, gg_activity, gg_excitation, gg_intensity,
  nrow = 4, labels = "auto", align = "hv"
)
```
