---
title: "Manuscript Figures & Tables"
subtitle: "Assessing Predictive Performance"
author: "J.P. Meagher"
date: today
format:
  pdf:
    keep-tex: true
    fig-width: 6
    fig-crop: false
execute:
  echo: false
  warning: false
  message: false
knitr:
  opts_chunk:
    crop: null
    dev: "pdf"
---

```{r setup}
library(onlineMessageboardActivity)
library(dplyr)
library(ggplot2)
library(xtable)
library(reshape2)
library(latex2exp)
library(magrittr)
library(tidyr)

source(here::here("code_modules/plotting.R"))
```

```{r load_data}
prediction_summary <- readRDS(here::here("predictions/prediction_summary.RDS"))
sim_sizes <- readRDS(here::here("predictions/simulated_sizes.RDS"))
# sim_sizes is a list over 5 models, each containing a list over 7 observation horizons,
# each containing a 100 x 3716 matrix of simulated discussion sizes for the 3716 test discussions.

# observation intervals used in assess_predictive_performance.R
s <- c(0, 2^(0:5))

# model labels for plotting
model_labels <- c(
  TeX("$M_1$"), TeX("$M_2$"), TeX("$M_3$"), TeX("$M_4$"), TeX("$M_5$")
)
```

# Expected Log Predictive Density

```{r lpd_prediction_table, results='asis'}
prediction_summary |>
  as.data.frame() |>
  select(starts_with("elpd")) |>
  mutate(across(everything(), ~ .x - prediction_summary$elpd_M4)) |>
  summarise(across(everything(), ~ {
    n <- sum(!is.na(.x))
    m <- sum(.x)
    se <- sd(.x) * sqrt(n)
    paste0(
      sprintf("%.1f", m), " (", sprintf("%.1f", se), ")"
    )
  })) |>
  set_colnames(
    c(
      "$\\mathcal{M}_1$", "$\\mathcal{M}_2$", "$\\mathcal{M}_3$",
      "$\\mathcal{M}_4$", "$\\mathcal{M}_5$"
    )
  ) |>
  set_rownames("$\\Delta \\widehat{\\operatorname{lpd}}_{l4}$") |>
  xtable(
    label = "tab:lpd",
    caption = "The difference (standard error) in log cluster-wise predictive density on $\\boldsymbol y_{\\operatorname{test}}$ between each model and $\\mathcal M_4$.",
    align = "|c|c|c|c|c|c|"
  ) |>
  print.xtable(
    include.rownames = TRUE,
    include.colnames = TRUE,
    type = "latex",
    sanitize.text.function = function(x) x,
    comment = FALSE
  )
```

# CRPS Skill Scores

```{r baseline_crps}
train_size <- train_df |>
  group_by(discussion) |>
  summarise(size = length(discussion)) |>
  pull(size)

baseline_crps <- sapply(
  prediction_summary$size,
  compute_pwm_crps,
  pred = train_size,
  perform_checks = FALSE
) |>
  mean()
```

```{r compute_crps}
crps_df <- prediction_summary |>
  select(contains("crps")) |>
  summarise(across(
    everything(),
    list(
      mean = mean,
      se = \(x) sd(x) / sqrt(length(x))
    )
  )) |>
  reshape2::melt() |>
  mutate(variable = as.character(variable)) |>
  mutate(
    model_s = sapply(strsplit(variable, "_"), \(x) x[2]),
    stat = sapply(strsplit(variable, "_"), \(x) x[3]),
    model = sapply(strsplit(model_s, ".", fixed = TRUE), \(x) x[1]),
    s = as.numeric(sapply(strsplit(model_s, ".", fixed = TRUE), \(x) x[2])),
    s = c(0, 2^(0:5))[s]
  )

crps_df <- crps_df |>
  filter(stat == "mean") |>
  mutate(mean = value) |>
  cbind(
    se = crps_df |>
      filter(stat == "se") |>
      pull(value)
  ) |>
  filter(s < 10)
```

```{r crps_assessment, fig.height=3, fig.width=7, fig.cap="CRPS skill scores for predicting discussion size relative to the empirical distribution baseline."}
crps_df |>
  mutate(
    lower = 1 - ((mean - se) / baseline_crps),
    upper = 1 - ((mean + se) / baseline_crps),
    mean = 1 - (mean / baseline_crps)
  ) |>
  ggplot() +
  geom_point(aes(x = s, y = mean, color = model)) +
  geom_line(aes(x = s, y = mean, color = model)) +
  geom_errorbar(
    aes(x = s, ymin = lower, ymax = upper, color = model),
    width = 0.25
  ) +
  scale_color_viridis_d(labels = model_labels) +
  my_theme +
  labs(
    title = "Predicting discussion size",
    y = "Skill score",
    x = "Observation interval (hours after immigrant seed)",
    color = NULL
  )
```

# Empirical Coverage Probabilities

For each model $\mathcal{M}_l$ and observation horizon $s$, we compute the empirical coverage probability at nominal levels $1 - \alpha$ for $\alpha \in \{0.20, 0.10, 0.05\}$. For each discussion $i$, a $(1 - \alpha)$ highest density region (HDR) is constructed from the 100 posterior predictive simulations of $n_i$. The HDR is the shortest interval $[l, u]$ containing at least $(1-\alpha)$ of the simulated values. For the right-skewed count distributions typical of discussion sizes, the HDR is generally anchored at the mode (often 1), producing sharper intervals than the conventional equal-tailed approach.

The empirical coverage is the proportion of discussions for which the observed $n_i$ falls within the HDR.

```{r compute_coverage}
nominal_levels <- c(0.80, 0.90, 0.95)
model_names <- c("M1", "M2", "M3", "M4", "M5")

compute_hdr_coverage <- function(sim_mat, observed, level) {
  # HDInterval::hdi on a matrix returns a 2-row matrix (lower, upper)
  # with one column per input column, i.e. one column per discussion.
  hdr <- HDInterval::hdi(sim_mat, credMass = level)
  mean(observed >= hdr[1, ] & observed <= hdr[2, ])
}

coverage_df <- expand.grid(
  model = model_names,
  s_idx = seq_along(s),
  nominal = nominal_levels,
  stringsAsFactors = FALSE
) |>
  mutate(
    s_val = s[s_idx],
    empirical = purrr::pmap_dbl(
      list(model, s_idx, nominal),
      function(m, h, lvl) {
        compute_hdr_coverage(
          sim_sizes[[m]][[h]],
          prediction_summary$size,
          lvl
        )
      }
    )
  )
```

```{r coverage_table, results='asis'}
# Coverage at s = 0 (no observation) for all models
coverage_df |>
  filter(s_val == 0) |>
  mutate(
    nominal_pct = paste0(nominal * 100, "\\%"),
    empirical_pct = sprintf("%.1f\\%%", empirical * 100),
    model_label = paste0("$\\mathcal{M}_", gsub("M", "", model), "$")
  ) |>
  select(model_label, nominal_pct, empirical_pct) |>
  pivot_wider(
    names_from = model_label,
    values_from = empirical_pct
  ) |>
  set_colnames(c(
    "Nominal",
    "$\\mathcal{M}_1$", "$\\mathcal{M}_2$", "$\\mathcal{M}_3$",
    "$\\mathcal{M}_4$", "$\\mathcal{M}_5$"
  )) |>
  xtable(
    caption = "Empirical coverage probabilities based on highest density regions for predicted discussion size $n_i$ at $s = 0$ (prediction from the immigrant seed only). Each HDR is constructed from 100 posterior predictive simulations. Well-calibrated prediction regions should have empirical coverage close to the nominal level.",
    label = "tab:coverage_s0",
    align = "|l|c|c|c|c|c|c|"
  ) |>
  print.xtable(
    include.rownames = FALSE,
    type = "latex",
    sanitize.text.function = function(x) x,
    comment = FALSE
  )
```

```{r coverage_by_horizon, fig.height=4, fig.width=7, fig.cap="Empirical coverage probabilities of highest density regions as a function of observation horizon for 90\\% prediction regions. The dashed line indicates the nominal 90\\% level."}
coverage_df |>
  filter(nominal == 0.90, s_val < 10) |>
  ggplot(aes(x = s_val, y = empirical, color = model)) +
  geom_hline(yintercept = 0.90, lty = 2) +
  geom_point() +
  geom_line() +
  scale_color_viridis_d(labels = model_labels) +
  scale_y_continuous(
    limits = c(0.5, 1),
    labels = scales::percent_format()
  ) +
  my_theme +
  labs(
    title = "Empirical coverage of 90% highest density regions",
    x = "Observation interval (hours after immigrant seed)",
    y = "Empirical coverage",
    color = NULL
  )
```
