---
title: "Manuscript Figures & Tables"
subtitle: "Goodness of Fit"
author: "J.P. Meagher"
date: today
format:
  pdf:
    keep-tex: true
    fig-width: 6
    fig-crop: false
execute:
  echo: false
  warning: false
  message: false
knitr:
  opts_chunk:
    crop: null
    dev: "pdf"
---

```{r setup}
library(onlineMessageboardActivity)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(latex2exp)
library(magrittr)
library(reshape2)
library(stringr)

source(here::here("code_modules/plotting.R"))
source(here::here("code_modules/helpers.R"))
```

## Figure 7

This figure presents an analysis of the goodness of fit for each model to the set of all discussions seeded during the training interval.

```{r load_data}
discussion_size <- readRDS(here::here("predictions/discussion_size.RDS"))

# Select only those discussions included in the training set
discussion_size <- discussion_size |>
  filter(discussion %in% train_df$discussion)
```

We first compute bootstrapped estimates of the KS test statistic comparing predicted cluster sizes from each model to the empirical data.

```{r ks_test_statistic}
ks_boot <- replicate(1000, {
  tmp_df <- discussion_size |>
    slice_sample(n = 2017, replace = TRUE)

  tmp_df |>
    select(starts_with("M")) |>
    apply(2, function(x) {
      ks.test(tmp_df$n, x)$statistic
    })
})
```

```{r ks_check, fig.height=3, fig.width=6}
gg_ks <- ks_boot |>
  t() |>
  data.frame() |>
  melt() |>
  ggplot() +
  geom_histogram(
    aes(x = value, y = after_stat(density), fill = variable),
    position = "identity", alpha = 0.5, binwidth = 0.0025
  ) +
  scale_fill_viridis_d(
    labels = c(TeX("$M_1$"), TeX("$M_2$"), TeX("$M_3$"), TeX("$M_4$"), TeX("$M_5$"))
  ) +
  my_theme +
  theme(legend.position = "bottom") +
  labs(
    title = "Bootstrap estimates",
    x = "Kolmogorov-Smirnov test statistic",
    y = "Density",
    fill = NULL
  )
```

We then compare the mean discussion size for 3-hour blocks during the day coomparing our the simplest generative model and the full generative model to the empirical data.

```{r time_evolution_data}
ds_df <- discussion_size |>
  mutate(two_hour_bin = floor(tod / 3) * 3) |>
  select(two_hour_bin, n, dplyr::starts_with("M")) |>
  group_by(two_hour_bin) |>
  summarise(
    across(everything(), list(mean = mean, se = se)),
    .groups = "drop"
  ) |>
  mutate(time = two_hour_bin + 1.5)
```

```{r time_evolution_discussion_size, fig.height=3, fig.width=6}
time_evo_df <- ds_df |>
  select(time, ends_with("mean")) |>
  melt(id.vars = "time") |>
  mutate(mean = value) |>
  select(-value) |>
  cbind(
    ds_df |>
      select(time, ends_with("se")) |>
      melt(id.vars = "time") |>
      select(se = value)
  ) |> 
  mutate(variable = ifelse(variable == "n_mean", "data", str_replace(variable, "_mean", "")))

gg_time_evo <- time_evo_df |>
  ggplot() +
  geom_hline(yintercept = mean(discussion_size$n), lty = 2) +
  geom_errorbar(
    data = filter(time_evo_df, variable == "data"),
    aes(x = time, ymin = mean - 2 * se, ymax = mean + 2 * se)
  ) +
  geom_point(
    data = filter(time_evo_df, variable == "data"),
    aes(x = time, y = mean)
  ) +
  geom_line(
    data = filter(time_evo_df, variable %in% c("M1", "M4")),
    aes(x = time, y = mean, color = variable)
  ) +
  my_theme +
  scale_color_manual(
    values = c(viridisLite::viridis(5)[1], viridisLite::viridis(5)[4]),
    labels = expression(M[1], M[4])
  ) +
  theme(legend.position = "none") +
  scale_x_continuous(breaks = c(0, 6, 12, 18, 24)) +
  labs(
    title = "The time evolution of mean discussion size",
    subtitle = expression(M[1] * " and " * M[4] * " compared to empirical data"),
    x = "Hour of the day",
    y = "Mean discussion size",
    color = NULL
  )

gg_time_evo
```

Finally, we produce quantile-quantile plots comparing the distribution of discussion sizes for each model to the empirical distribution.

```{r qq_data}
qq_df <- discussion_size |>
  select(-time, -t, -tod, -discussion) |>
  melt(id.vars = "n") |>
  group_by(variable) |>
  mutate(n = sort(n), value = sort(value)) |>
  ungroup()
```

```{r qq_plots, fig.height=3, fig.width=6}
gg_qq <- qq_df |>
  ggplot() +
  geom_abline(slope = 1, intercept = 0) +
  geom_point(aes(x = n, y = value, color = variable)) +
  facet_wrap(vars(variable), nrow = 1) +
  scale_x_log10() +
  scale_y_log10() +
  my_theme +
  theme(
    strip.background = element_blank(),
    strip.text.x = element_blank(),
    legend.position = "bottom"
  ) +
  scale_color_viridis_d(
    labels = c(TeX("$M_1$"), TeX("$M_2$"), TeX("$M_3$"), TeX("$M_4$"), TeX("$M_5$"))
  ) +
  labs(
    title = "Q-Q plots",
    x = "Empirical discussion sizes",
    y = "Simulated discussion sizes",
    color = NULL
  )
```

```{r 04_analysisAndResults_fig-gof, fig.height=6, fig.width=6, dev="png", dpi=120}
# Note: a PDF version of this plot results in an overly large file,
# so this figure is saved as a PNG.
gg_cm_gof <- ggarrange(
  gg_ks, gg_qq,
  common.legend = TRUE, labels = "auto", align = "hv"
)

ggarrange(gg_cm_gof, gg_time_evo, nrow = 2, labels = list("", "c"))
```

```{r ks_comparisons, eval=FALSE}
# Code for computations that inform discussion in the manuscript
mean(ks_boot[1, ] > ks_boot[4, ])
mean(ks_boot[1, ] > ks_boot[5, ])

mean(ks_boot[2, ] > ks_boot[4, ])
mean(ks_boot[2, ] > ks_boot[5, ])

mean(ks_boot[3, ] > ks_boot[4, ])
mean(ks_boot[3, ] > ks_boot[5, ])

mean(ks_boot[4, ] > ks_boot[5, ])
```
